\fullpagehumor[{Radio Yerevan Jokes}]{

    \textbf{Q: Radio Yerevan was asked: \QENOpen{}Capitalism is the exploitation of man by man. What about Socialism?\QENClose{}} \\  
    \textbf{A:} Radio Yerevan answered: \QENOpen{}Under Socialism it is exactly the other way around.\QENClose{}  

    \medskip  
    \textbf{Q: Is there freedom of speech in the Soviet Union the same as in the USA?} \\  
    \textbf{A:} In principle, yes. In the USA, you can stand in front of the Washington Monument and yell, \QENOpen{}Down with Reagan!\QENClose{} and you will not be punished. In the Soviet Union, you can stand in Red Square and yell, \QENOpen{}Down with Reagan!\QENClose{} and you will not be punished.  

    \medskip  
    \textbf{Q: What is chaos?} \\  
    \textbf{A:} We do not comment on national economics.  

    \medskip  
    \textbf{Q: Is it true that half of the members of the Central Committee are fools?} \\  
    \textbf{A:} What a crazy question. Half of the members of the Central Committee are not fools!  

    \medskip  
    \textbf{Q: Is it true that Ivan Ivanovich Ivanov from Moscow won a car in a lottery?} \\  
    \textbf{A:} In principle, yes. But: it wasn't Ivan Ivanovich Ivanov but Aleksander Aleksandrovich Aleksandrov; he is not from Moscow but from Odessa; it was not a car but a bicycle; and he didn’t win it—it was stolen from him.  

    \medskip  
    \textbf{Q: Why is our government not in a hurry to land our men on the moon?} \\  
    \textbf{A:} What if they refuse to return?  

    \medskip  
    \textbf{Q: We are told that communism is already visible on the horizon. What then is a horizon?} \\  
    \textbf{A:} An imaginary line that moves farther away each time you approach it.  

    \medskip  
    \textbf{Q: Is it true that conditions in our labour camps are excellent?} \\  
    \textbf{A:} In principle, yes. Five years ago, one of our listeners was sceptical, so he was sent to investigate. He must have liked it so much that he hasn’t returned yet.  

    \medskip  
    \textbf{Q: What is the most beautiful city in the Soviet Union?} \\  
    \textbf{A:} Yerevan, obviously.\\
    \textbf{Q: How many nuclear bombs will it take to destroy Yerevan?} \\  
    \textbf{A:} Baku is also a beautiful city.  

    \medskip  
    \textbf{Q: Is it true that the United States is on the edge of a precipice?} \\  
    \textbf{A:} True, but we are a step farther than them.  
}
