\chapter*{Bibliography}
\addcontentsline{toc}{chapter}{Bibliography}
\markboth{BIBLIOGRAPHY}{BIBLIOGRAPHY}

\begingroup
\raggedright
\emergencystretch=2em
\setlength{\parindent}{0pt}

\subsection*{Scholarly Works}
\vspace{0.3em}
\begin{list}{}{%
  \setlength{\leftmargin}{2em}%
  \setlength{\itemindent}{-2em}%
  \setlength{\itemsep}{0.35\baselineskip}%
  \setlength{\parsep}{0pt}%
  \setlength{\topsep}{0.5em}}
\small
\item (1914). `Christmas at the Front/An Interrupted Truce'. \textit{Morning Post}. \hfill {\scriptsize [Ch.~38]}
\item Albert Einstein (1905). `Zur Elektrodynamik bewegter Körper'. \textit{Annalen der Physik} \textbf{322}(10), 891–921. \hfill {\scriptsize [Ch.~1, 14, 19]}
\item Albert Einstein (1907). Über das Relativitätsprinzip und die aus demselben gezogenen Folgerungen. Jahrbuch der Radioaktivität und Elektronik.
\item Albert Einstein (1916). Die Grundlage der allgemeinen Relativitätstheorie. Annalen der Physik. \hfill {\scriptsize [Ch.~6]}
\item Albert Einstein \& Nathan Rosen (1935). The Particle Problem in the General Theory of Relativity. Physical Review.
\item Albrecht Dürer (1525). \textit{Underweysung der messung mit dem zirckel und richtscheyt in Linien ebnen unnd gantzen corporen}. Hieronymus Andreae, Nuremberg. \hfill {\scriptsize [Ch.~34]}
\item Alexandru Scorpan (2005). \textit{The Wild World of 4-Manifolds}. American Mathematical Society. \hfill {\scriptsize [Ch.~24]}
\item Allan Gibbard (1973). `Manipulation of Voting Schemes: A General Result'. \textit{Econometrica 41(4)}. \hfill {\scriptsize [Ch.~9]}
\item Anaxagoras, c. 450 BCE. Fragment DK 59 A42, reported by various ancient sources including Diogenes Laërtius. \hfill {\scriptsize [Ch.~10]}
\item Annie Jump Cannon, Edward C. Pickering (1918-1924). \textit{Henry Draper Catalog}. Harvard College Observatory. \hfill {\scriptsize [Ch.~27]}
\item Anonymous (1915). `Letter from a doctor attached to the Rifle Brigade'. \textit{The Times}. \hfill {\scriptsize [Ch.~39]}
\item Archimedes (c. 250 BCE). \textit{The Sand Reckoner}. \hfill {\scriptsize [Ch.~17]}
\item Ashlock \& Tillotson (1993). An automaton for game-theoretic representation. Proceedings of the 2nd Annual Conference on Evolutionary Programming.
\item Atila, A., Sukhomlinov, S. V., and M\"{u}ser, M. H. (2024). \textit{Phys. Rev. Lett.}, \textbf{133}, 236201. \hfill {\scriptsize [Ch.~42]}
\item Augustine of Hippo (c. 397). \textit{Confessions}. Book XI. \hfill {\scriptsize [Ch.~43]}
\item Axelrod (1984). The Evolution of Cooperation. Basic Books.
\item B. T. Draine (2003). `Interstellar Dust Grains'. \textit{Annual Review of Astronomy and Astrophysics}. \hfill {\scriptsize [Ch.~27]}
\item Babinet (1837). Mémoire sur les couleurs des réseaux. Comptes Rendus de l'Académie des Sciences.
\item Ballard (1981). Generalizing the Hough transform to detect arbitrary shapes. Pattern Recognition.
\item Banach, S. and Tarski, A. (1924). \textit{Fund. Math.} \textbf{6} \hfill {\scriptsize [Ch.~3]}
\item Bardeen, Carter \& Hawking (1973). The Four Laws of Black Hole Mechanics. Communications in Mathematical Physics.
\item Barkan, E., Biham, E., Keller, N. (2008). Instant ciphertext-only cryptanalysis of GSM encrypted communication. \textit{J. Cryptology} 21(3):392–429. \hfill {\scriptsize [Ch.~12]}
\item Baryshnikov, Y. (1997). Topological and discrete social choice: in a search of a theory. \textit{Social Choice and Welfare}, \textbf{14}(2), 199–209. \hfill {\scriptsize [Ch.~9]}
\item Bassham, Benson \& Calvin (1950). The path of carbon in photosynthesis. Journal of Biological Chemistry.
\item Benford (1938). The law of anomalous numbers. Proceedings of the American Philosophical Society.
\item Benjamin W. Fortson IV (2010). \textit{Indo-European Language and Culture: An Introduction}. Wiley-Blackwell. \hfill {\scriptsize [Ch.~5]}
\item Bennett, C. H. (1982). \textit{The Thermodynamics of Computation — A Review}. Int. J. Theor. Phys., 21, 905–940. \hfill {\scriptsize [Ch.~45]}
\item Bentley et al. (2008). \textit{Nature} 456:53–59. \hfill {\scriptsize [Ch.~40]}
\item Berkson (1946). Limitations of the application of fourfold table analysis to hospital data. Biometrics Bulletin.
\item Bernevig, B. A., Hughes, T. L., and Zhang, S.-C. (2006). Quantum Spin Hall Effect and Topological Phase Transition in HgTe Quantum Wells. \textit{Science}, \textbf{314}(5806), 1757–1761. \hfill {\scriptsize [Ch.~11]}
\item Bethe, H. A. (1939). Energy Production in Stars. \textit{Phys. Rev.}, \textbf{55}, 434–456. \hfill {\scriptsize [Ch.~10]}
\item Bloch (1929). Über die Quantenmechanik der Elektronen in Kristallgittern. Zeitschrift für Physik.
\item Bogoliubov (1947). On the theory of superfluidity. Journal of Physics (USSR).
\item Boltzmann (1877). Über die Beziehung zwischen dem zweiten Hauptsatze der mechanischen Wärmetheorie und der Wahrscheinlichkeitsrechnung. Wiener Berichte. \hfill {\scriptsize [Ch.~28, 34, 45]}
\item \textit{Book of Joshua}. Hebrew Bible. \hfill {\scriptsize [Ch.~38]}
\item Bopp (1816). Über das Conjugationssystem der Sanskritsprache in Vergleichung mit jenem der griechischen, lateinischen, persischen und germanischen Sprache. Andreäische Buchhandlung.
\item Boyle \& Smith (1970). Charge-Coupled Semiconductor Devices. Bell System Technical Journal.
\item Brascamp \& Lieb (1976). Best constants in Young's inequality, its converse, and its generalization to more than three functions. Advances in Mathematics.
\item C. J. Isham (1999). \textit{Modern Differential Geometry for Physicists}. World Scientific. \hfill {\scriptsize [Ch.~14]}
\item Cantor (1891). Über eine elementare Frage der Mannigfaltigkeitslehre. Jahresbericht der Deutschen Mathematiker-Vereinigung.
\item Captain Josef Sewald, 17th Bavarian Regiment. Quoted in Stanley Weintraub, \textit{Silent Night: The Story of the World War I Christmas Truce} (New York: Free Press, 2001). \hfill {\scriptsize [Ch.~38]}
\item Carroll, S. M. (2001). The cosmological constant. \textit{Living Rev. Relativ.}, 4, 1. \hfill {\scriptsize [Ch.~2]}
\item Casimir (1948). On the attraction between two perfectly conducting plates. Proceedings of the Royal Netherlands Academy of Arts and Sciences.
\item Chandrasekhar (1931). The Maximum Mass of Ideal White Dwarfs. The Astrophysical Journal.
\item Christenson, J. H., Cronin, J. W., Fitch, V. L., and Turlay, R. (1964). Evidence for the $2\pi$ Decay of the $K_2^0$ Meson. \textit{Phys. Rev. Lett.} \textbf{13}, 138–140. \hfill {\scriptsize [Ch.~19]}
\item Clausius (1850). Ueber die bewegende Kraft der Wärme. Annalen der Physik. \hfill {\scriptsize [Ch.~42]}
\item Clausius (1870). On a Mechanical Theorem Applicable to Heat. Philosophical Magazine. \hfill {\scriptsize [Ch.~42]}
\item Clayton, D. D. (1983). \textit{Principles of Stellar Evolution and Nucleosynthesis}. University of Chicago Press. \hfill {\scriptsize [Ch.~10]}
\item Clowe, D., Brada\v{c}, M., Gonzalez, A. H., Markevitch, M., Randall, S. W., Jones, C., and Zaritsky, D. (2006). A Direct Empirical Proof of the Existence of Dark Matter. \textit{ApJL}, \textbf{648}, L109-L113. \hfill {\scriptsize [Ch.~37]}
\item Coblentz, W.W. (1912). \textit{A Physical Study of the Firefly}. Carnegie Institution of Washington, Publ. 164. \hfill {\scriptsize [Ch.~25]}
\item Compton (1923). A Quantum Theory of the Scattering of X-rays by Light Elements. Physical Review.
\item Condorcet (1785). Essai sur l'application de l'analyze à la probabilité des décisions rendues à la pluralité des voix. Imprimerie Royale.
\item D. G. Saari (1994). \textit{Geometry of Voting}. Springer-Verlag. \hfill {\scriptsize [Ch.~9]}
\item Dahl \& Nygaard (1966). SIMULA — an ALGOL-based simulation language. Communications of the ACM.
\item Daniel C. Dennett (1987). \textit{The Intentional Stance}. MIT Press. \hfill {\scriptsize [Ch.~20]}
\item Daniel J. Bernstein (1999). `Bernstein v. United States Department of Justice'. \textit{9th Circuit, 176 F.3d 1132 (1999); withdrawn on rehearing en banc, 192 F.3d 1308}. \hfill {\scriptsize [Ch.~17]}
\item David Hilbert, "Über das Unendliche," \textit{Mathematische Annalen} 95 (1926), pp. 161–190. \hfill {\scriptsize [Ch.~1]}
\item David Hilbert, as quoted in Calvin C. Clawson, \textit{Mathematical Mysteries: The Beauty and Magic of Numbers} (New York: Basic Books, 1999), p. 258. \hfill {\scriptsize [Ch.~1]}
\item David J. Chalmers (1995). `Facing Up to the Problem of Consciousness'. \textit{Journal of Consciousness Studies}. \hfill {\scriptsize [Ch.~1, 50]}
\item David J. Chalmers (1996). \textit{The Conscious Mind: In Search of a Fundamental Theory}. Oxford University Press. \hfill {\scriptsize [Ch.~1, 20]}
\item Davis, R., Harmer, D. S., and Hoffman, K. C. (1968). Search for Neutrinos from the Sun. \textit{Phys. Rev. Lett.} \textbf{20}, 1205–1209. \hfill {\scriptsize [Ch.~10]}
\item de la Vall\'ee Poussin, C.-J. (1896). Recherches analytiques sur la th\'eorie des nombres premiers. Annales de la Soci\'et\'e Scientifique de Bruxelles.
\item Debye, P. (1923). Physikalische Eigenschaften von Lösungen und Theorie der Elektrolyte. \textit{Phys. Z.} 24:334–338. \hfill {\scriptsize [Ch.~31]}
\item Degrassi et al., \textit{JHEP} \textbf{08}, 098 (2012). \hfill {\scriptsize [Ch.~16]}
\item Diels \& Alder (1928). Synthesen in der hydroaromatischen Reihe. Justus Liebigs Annalen der Chemie.
\item Dirac, P. A. M. (1928). `The Quantum Theory of the Electron'. \textit{Proc. R. Soc. Lond. A} \textbf{117}, 610–624. \hfill {\scriptsize [Ch.~1]}
\item Dodelson, Scott (2003). \textit{Modern Cosmology}. Academic Press. \hfill {\scriptsize [Ch.~43]}
\item Donaldson (1983). An application of gauge theory to four-dimensional topology. Journal of Differential Geometry.
\item Du Junca, Étienne (1690-1705). `Registres des entrées et des sorties de la Bastille'. \textit{Bibliothèque de l'Arsenal, Paris}. \hfill {\scriptsize [Ch.~39]}
\item Dubois (1887). Ueber die Magnetisirungsconstante des Wassers und den absoluten Werth der Susceptibilität des Eisens. Annalen der Physik und Chemie.
\item Duda, R. O., and Hart, P. E. (1972). Use of the Hough Transformation to Detect Lines and Curves in Pictures. \textit{Commun. ACM} 15(1), 11–15. \hfill {\scriptsize [Ch.~41]}
\item Dunkel, J., Hilbert, S. (2014). \textit{Consistent Thermostatistics Forbids Negative Absolute Temperatures}. Nat. Phys., 10, 67. \hfill {\scriptsize [Ch.~28]}
\item Dyson (1949). The S-Matrix in Quantum Electrodynamics. Physical Review. \hfill {\scriptsize [Ch.~15]}
\item Dyson, L., Kleban, M., Susskind, L. (2002). Disturbing implications of a cosmological constant. \textit{JHEP} 0210:011. \hfill {\scriptsize [Ch.~34]}
\item Eccles \& Jordan (1919). A trigger relay utilizing three-electrode thermionic vacuum tubes. The Radio Review.
\item Ecclesiastes 1:2, Hebrew Bible. \hfill {\scriptsize [Ch.~34]}
\item Eddington (1931). The End of the World: from the Standpoint of Mathematical Physics. Nature. \hfill {\scriptsize [Ch.~34]}
\item Edward Kasner and James R. Newman (1940). \textit{Mathematics and the Imagination}. Simon \& Schuster. \hfill {\scriptsize [Ch.~17]}
\item Egan, G. (2018). `Superpermutations'. \textit{gregegan.net/SCIENCE/Superpermutations/}. \hfill {\scriptsize [Ch.~39]}
\item Ehrenfest, P. (1917). \textit{In what way does it become manifest in the fundamental laws of physics that space has three dimensions?} Proceedings of the Amsterdam Academy. \hfill {\scriptsize [Ch.~24]}
\item Eisenstein, D. J. et al. (2005). Detection of the Baryon Acoustic Peak. \textit{ApJ} \textbf{633}, 560–574. \hfill {\scriptsize [Ch.~2]}
\item Elias (1955). Coding for noisy channels. IRE Convention Record.
\item Englert \& Brout (1964). Broken Symmetry and the Mass of Gauge Vector Mesons. Physical Review Letters.
\item Erdős, P. (1934). `A Theorem of Sylvester and Schur'. \textit{J. London Math. Soc.} \textbf{9}, 282–288. \hfill {\scriptsize [Ch.~8, 9]}
\item Eric Hobsbawm, Terence Ranger (eds.) (1983). \textit{The Invention of Tradition}. Cambridge University Press. \hfill {\scriptsize [Ch.~38]}
\item Eugene Wigner, "The Unreasonable Effectiveness of Mathematics in the Natural Sciences," \textit{Communications in Pure and Applied Mathematics} 13, no. 1 (February 1960), pp. 1–14. \hfill {\scriptsize [Ch.~24]}
\item Exodus 10:21, Hebrew Bible. \hfill {\scriptsize [Ch.~37]}
\item Farquhar, G. D., von Caemmerer, S., Berry, J. A. (1980). A biochemical model of photosynthetic $\mathrm{CO}_2$ assimilation in C$_3$ leaves. \textit{Planta}, \textbf{149}, 78–90. \hfill {\scriptsize [Ch.~35]}
\item Feigenbaum, M. J. (1978). Quantitative universality for a class of nonlinear transformations. \textit{Journal of Statistical Physics}, 19, 25–52. \hfill {\scriptsize [Ch.~48]}
\item Feldman, W. M. (1931). \textit{Rabbinical Mathematics and Astronomy}. M. L. Cailingold. \hfill {\scriptsize [Ch.~26]}
\item Ferro, M. (1990). \textit{The Great War 1914–1918}. Routledge. \hfill {\scriptsize [Ch.~38]}
\item Festinger (1957). A Theory of Cognitive Dissonance. Stanford University Press.
\item Feynman, R. P., Leighton, R. B., and Sands, M. (1964). \textit{The Feynman Lectures on Physics}. Addison-Wesley. \hfill {\scriptsize [Ch.~4]}
\item Fick (1855). Ueber Diffusion. Annalen der Physik.
\item Finkelstein, A. (1987). Water Movement Through Lipid Bilayers, Pores, and Plasma Membranes. \textit{Wiley}. \hfill {\scriptsize [Ch.~31]}
\item Frank Richards (1933). \textit{Old Soldiers Never Die}. Faber and Faber Limited. \hfill {\scriptsize [Ch.~38]}
\item Friedmann (1922). Über die Krümmung des Raumes. Zeitschrift für Physik. \hfill {\scriptsize [Ch.~2]}
\item G. H. Hardy (1940). \textit{A Mathematician's Apology}. Cambridge University Press. \hfill {\scriptsize [Ch.~8]}
\item G. Udny Yule (1911). \textit{An Introduction to the Theory of Statistics}. C. Griffin and Company, Limited. \hfill {\scriptsize [Ch.~30]}
\item Gamow (1928). Zur Quantentheorie des Atomkernes. Zeitschrift für Physik.
\item Gauss (1809). Theoria motus corporum coelestium in sectionibus conicis solem ambientium. F. Perthes and I. H. Besser. \hfill {\scriptsize [Ch.~33]}
\item Geroch, R. (1967). \textit{J. Math. Phys.}, \textbf{8}. \hfill {\scriptsize [Ch.~14]}
\item Gibbs (1876). On the Equilibrium of Heterogeneous Substances. Transactions of the Connecticut Academy of Arts and Sciences.
\item Golomb, S. W. (1955). Sequences with Randomness Properties. Glenn L. Martin Company.
\item Goroff, M. H., and Sagnotti, A. (1986). The ultraviolet behavior of Einstein gravity. \textit{Nuclear Physics B}. \hfill {\scriptsize [Ch.~36]}
\item Gottfried Wilhelm Leibniz (1686). \textit{Discourse on Metaphysics}. \hfill {\scriptsize [Ch.~24]}
\item Grimm, J. (1822). \textit{Deutsche Grammatik}. Göttingen: Dieterich. \hfill {\scriptsize [Ch.~5]}
\item Gross, D. J. \& Wilczek, F. (1973). Ultraviolet Behavior of Non-Abelian Gauge Theories. Physical Review Letters.
\item Gödel, K. (1931). `Über formal unentscheidbare Sätze der Principia Mathematica und verwandter Systeme I'. \textit{Monatshefte für Mathematik und Physik} \textbf{38}, 173–198. \hfill {\scriptsize [Ch.~3, 21]}
\item Gödel, K. (1938). The Consistency of the Axiom of Choice and of the Generalized Continuum-Hypothesis. \textit{Proc. Natl. Acad. Sci.} \textbf{24}(12), 556–557. \hfill {\scriptsize [Ch.~3]}
\item H. Furstenberg (1955). `On the Infinitude of Primes'. \textit{American Mathematical Monthly}. \hfill {\scriptsize [Ch.~8]}
\item H. Poincaré (1890). `Sur le problème des trois corps et les équations de la dynamique'. \textit{Acta Mathematica}. \hfill {\scriptsize [Ch.~48]}
\item Hadamard, J. (1896). Sur la distribution des z\'eros de la fonction $\zeta(s)$ et ses cons\'equences arithm\'etiques. Bulletin de la Soci\'et\'e Math\'ematique de France.
\item Harvey, E. Newton (1920). \textit{The Nature of Animal Light}. J.B. Lippincott Company. \hfill {\scriptsize [Ch.~17]}
\item Harvey, E.N., and Stevens, K.L. (1928). The brightness of the light of the West Indian elaterid beetle, \textit{Pyrophorus}. \textit{J. Gen. Physiol.}, 12, 269–272. \hfill {\scriptsize [Ch.~25]}
\item Hegel, Georg Wilhelm Friedrich (1807). \textit{Phänomenologie des Geistes}. Joseph Anton Goebhardt, Bamberg. \hfill {\scriptsize [Ch.~50]}
\item Heisenberg, W. (1927). `Über den anschaulichen Inhalt der quantentheoretischen Kinematik und Mechanik'. \textit{Z. Phys.} \textbf{43}, 172–198. \hfill {\scriptsize [Ch.~1]}
\item Henri Poincaré, \textit{Calcul des Probabilités} (Paris: Gauthier-Villars, 1912), preface. \hfill {\scriptsize [Ch.~7]}
\item Henri Poincaré, \textit{La Science et l'Hypothèse} (Paris: Flammarion, 1902). \hfill {\scriptsize [Ch.~7]}
\item Henri Poincaré, \textit{Science et Méthode} (Paris: Flammarion, 1908). \hfill {\scriptsize [Ch.~7]}
\item Hermann Weyl (1918). \textit{Raum, Zeit, Materie: Vorlesungen über allgemeine Relativitätstheorie}. Springer, Berlin. \hfill {\scriptsize [Ch.~10]}
\item Hermann Weyl, "Invariants," \textit{Duke Mathematical Journal} 5 (1939), pp. 489–502. \hfill {\scriptsize [Ch.~11]}
\item Higgs, P. W. (1964). Broken Symmetries and the Masses of Gauge Bosons. Physical Review Letters.
\item Houston, R. (2014). \textit{Obvious Does Not Imply True: The Minimal Superpermutation Conjecture Is False}. arXiv:1408.5108. \hfill {\scriptsize [Ch.~39]}
\item Hubble, E. (1929). A Relation between Distance and Radial Velocity among Extra-Galactic Nebulae. Proceedings of the National Academy of Sciences. \hfill {\scriptsize [Ch.~2, 27, 43]}
\item Hugo Rossi, "Mathematics Is an Edifice, Not a Toolbox," \textit{Notices of the AMS} 43, no. 10 (October 1996), p. 1108. \hfill {\scriptsize [Ch.~21]}
\item Huw Price, "Retrieving the Baby: Reichenbach's Principle, Bell Locality, and Selection Bias," arXiv:2602.16985 (19 February 2026), Section 2, opening sentence. \hfill {\scriptsize [Ch.~30]}
\item Huygens, C. (1690). Trait\'e de la Lumi\`ere. Pieter van der Aa. \hfill {\scriptsize [Ch.~13, 24]}
\item Hyslop, L. A. et al. (2016). Towards clinical application of pronuclear transfer. \textit{Nature} \textbf{534}, 383–386. \hfill {\scriptsize [Ch.~49]}
\item International Bureau of Weights and Measures (BIPM) (2019). \textit{The International System of Units (SI)}. BIPM. \hfill {\scriptsize [Ch.~26]}
\item International Meridian Conference (1884). International Conference Held at Washington for the Purpose of Fixing a Prime Meridian and a Universal Day. Gibson Bros. \hfill {\scriptsize [Ch.~26]}
\item Isaac Newton, \textit{Principia Mathematica}, 2nd edition (1713), Definition VIII, Scholium. \hfill {\scriptsize [Ch.~6]}
\item Ives, H.E., and Coblentz, W.W. (1924). Photometric studies of luminous insects. \textit{J. Opt. Soc. Am.}, 9(3), 217–236. \hfill {\scriptsize [Ch.~25]}
\item J.D. Bernal, lecture at Birkbeck College, University of London (1960). Published in Alan L. Mackay and Maurice Ebison (eds.), \textit{The Harvest of a Quiet Eye} (London: Institute of Physics, 1977), p. 7. Original wording: "All that glisters may not be gold, but at least it contains free electrons." \hfill {\scriptsize [Ch.~1]}
\item James Clerk Maxwell, review in \textit{Nature} 17 (7 February 1878), p. 278. \hfill {\scriptsize [Ch.~34]}
\item James D. Watson and Francis H.C. Crick, "Molecular Structure of Nucleic Acids," \textit{Nature} 171 (25 April 1953), pp. 737–738. \hfill {\scriptsize [Ch.~4]}
\item James R. Newman (1956). \textit{The World of Mathematics}. Simon and Schuster. \hfill {\scriptsize [Ch.~4]}
\item Jinek, M. et al. (2012). A Programmable Dual-RNA–Guided DNA Endonuclease in Adaptive Bacterial Immunity. \textit{Science} \textbf{337}, 816–821. \hfill {\scriptsize [Ch.~49]}
\item John Archibald Wheeler. Earliest published version: "Space tells matter how to move; matter tells space how to curve," in C.W. Misner, K.S. Thorne, and J.A. Wheeler, \textit{Gravitation} (San Francisco: W.H. Freeman, 1973), p. 5. The "spacetime" variant appears in J.A. Wheeler and K. Ford, \textit{Geons, Black Holes, and Quantum Foam} (New York: Norton, 1998). \hfill {\scriptsize [Ch.~6, 14]}
\item John C. Baez, Javier P. Muniain (1994). \textit{Gauge Fields, Knots and Gravity}. World Scientific Publishing Company. \hfill {\scriptsize [Ch.~4]}
\item John Keegan (1999). \textit{The First World War}. A. Knopf. \hfill {\scriptsize [Ch.~38]}
\item John von Neumann, as recounted by Enrico Fermi to Freeman Dyson, 1953. Documented in Freeman Dyson, "A Meeting with Enrico Fermi," \textit{Nature} 427 (22 January 2004), p. 297. \hfill {\scriptsize [Ch.~1]}
\item Jürgen Mayer, Khaled Khairy, and Jonathon Howard, "Drawing an Elephant with Four Complex Parameters," \textit{American Journal of Physics} 78, no. 6 (June 2010), pp. 648–649. \hfill {\scriptsize [Ch.~41]}
\item Kahneman, D. \& Tversky, A. (1979). Prospect Theory: An Analysis of Decision under Risk. Econometrica.
\item Kane, C. L. and Mele, E. J. (2005). $\mathbb{Z}_2$ Topological Order and the Quantum Spin Hall Effect. \textit{Physical Review Letters}, \textbf{95}(14), 146802. \hfill {\scriptsize [Ch.~11]}
\item Kant, Immanuel (1781). \textit{Critique of Pure Reason}. Johann Friedrich Hartknoch, Riga. \hfill {\scriptsize [Ch.~50]}
\item Kasianowicz, J. J., Brandin, E., Branton, D., \& Deamer, D. W. (1996). Characterization of individual polynucleotide molecules using a membrane channel. Proceedings of the National Academy of Sciences.
\item Katori, H. (1999). Optical lattice clocks. Proceedings of the 6th Symposium on Frequency Standards and Metrology.
\item Kenneth J. Arrow (1963). \textit{Social Choice and Individual Values}. John Wiley \& Sons. \hfill {\scriptsize [Ch.~9]}
\item Kerr, R. P. (1963). `Gravitational Field of a Spinning Mass as an Example of Algebraically Special Metrics'. \textit{Phys. Rev. Lett.} \textbf{11}, 237–238. \hfill {\scriptsize [Ch.~23]}
\item Keynes, J. M. (1921). A Treatise on Probability. Macmillan and Co.
\item Kiefer, C. (2012). \textit{Quantum Gravity} (3rd ed.). Oxford University Press. \hfill {\scriptsize [Ch.~36]}
\item Klein, F. (1882). \"Uber Riemann's Theorie der Algebraischen Functionen und ihrer Integrale. B. G. Teubner. \hfill {\scriptsize [Ch.~14]}
\item Knuth, D. E. (1974). \textit{Surreal Numbers}. Addison-Wesley.
\item Kolb, E. W., and Turner, M. S. (1990). \textit{The Early Universe}. Addison-Wesley. \hfill {\scriptsize [Ch.~43]}
\item Kolmogorov, A. N. (1933). Grundbegriffe der Wahrscheinlichkeitsrechnung. Springer. \hfill {\scriptsize [Ch.~15]}
\item Kolmogorov, A. N. (1954). On Conservation of Conditionally Periodic Motions for a Small Change in Hamilton's Function. Doklady Akademii Nauk SSSR.
\item Kramers, H. A. (1930). Th\'eorie g\'en\'erale de la rotation paramagn\'etique dans les cristaux. Proceedings of the Amsterdam Academy.
\item Kruskal, M. D. (1960). Maximal extension of Schwarzschild metric. Physical Review
\item Landauer, R. (1961). \textit{Irreversibility and Heat Generation in the Computing Process}. IBM J. Res. Dev., 5(3), 183–191. \hfill {\scriptsize [Ch.~45]}
\item Langevin, P. (1911). L'évolution de l'espace et du temps. Scientia
\item Laplace, P. S. (1799). Traité de Mécanique Céleste. J.B.M. Duprat
\item Laplace, P. S. (1812). Théorie analytique des probabilités. Ve. Courcier \hfill {\scriptsize [Ch.~33]}
\item Laplace, P. S. (1814). Essai philosophique sur les probabilités. Courcier \hfill {\scriptsize [Ch.~48]}
\item Latała, R., Matlak, D. (2017). \textit{Royen's Proof of the Gaussian Correlation Inequality}. In: Israel Seminar (GAFA) 2014–2016. Springer. \hfill {\scriptsize [Ch.~33]}
\item Lawvere, F. W. (1969). Adjointness in foundations. Dialectica
\item Lenard, A., and Dyson, F. J. (1967). Stability of matter I, II. \textit{Journal of Mathematical Physics}. \hfill {\scriptsize [Ch.~24]}
\item Lense, J. \& Thirring, H. (1918). Über den Einfluß der Eigenrotation der Zentralkörper auf die Bewegung der Planeten und Monde nach der Einsteinschen Gravitationstheorie. Physikalische Zeitschrift
\item Leonardo da Vinci, c. 1490s–1500s. From his notebooks. Various translations exist. \hfill {\scriptsize [Ch.~5]}
\item Leopold Flatto (2009). \textit{Poncelet's Theorem}. American Mathematical Society. \hfill {\scriptsize [Ch.~7]}
\item Lincoln Taiz, Eduardo Zeiger, Ian Max Møller, Angus Murphy (2015). \textit{Plant Physiology and Development, Sixth Edition}. Sinauer Associates, Inc. \hfill {\scriptsize [Ch.~35]}
\item Liouville, J. (1838). Note sur la théorie de la variation des constantes arbitraires. Journal de Mathématiques Pures et Appliquées
\item Lloyd, S. P. (1965). Least squares quantization in PCM. Bell Laboratories Technical Memorandum \hfill {\scriptsize [Ch.~25]}
\item Lohmann, K. (1929). Über die Pyrophosphatfraktion im Muskel. Naturwissenschaften
\item Lopez, J. V., Yuhki, N., Masuda, R., Modi, W., and O'Brien, S. J. (1994). Numt, a recent transfer and tandem amplification of mitochondrial DNA to the nuclear genome of the domestic cat. \textit{J. Mol. Evol.} \textbf{39}, 174–190. \hfill {\scriptsize [Ch.~49]}
\item Lorenz, E. N. (1963). Deterministic nonperiodic flow. \textit{Journal of the Atmospheric Sciences}, 20(2), 130–141. \hfill {\scriptsize [Ch.~48]}
\item Ludlow, A. D. et al. (2015). \textit{Optical atomic clocks}. Rev. Mod. Phys. 87(2), 637–701. \hfill {\scriptsize [Ch.~24, 32]}
\item Lugaresi, E. et al. (1986). Fatal familial insomnia and dysautonomia with selective degeneration of thalamic nuclei. \textit{N. Engl. J. Med.} \textbf{315}, 997–1003. \hfill {\scriptsize [Ch.~50]}
\item Lyapunov, A. M. (1892). The general problem of the stability of motion. Kharkov Mathematical Society
\item Lynn F. Faust, Paul A. Weston (2009). `Degree-Day Prediction of Adult Emergence of Photinus carolinus (Coleoptera: Lampyridae)'. \textit{Environmental Entomology}. \hfill {\scriptsize [Ch.~25]}
\item Lynn Frierson Faust (2017). \textit{Fireflies, Glow-Worms, and Lightning Bugs}. University of Georgia Press. \hfill {\scriptsize [Ch.~25]}
\item M.C. Escher, "Approaches to Infinity," posthumously published essay (1973). Anthologized in Doris Schatschneider and Wallace Walker, eds. Full quote continues: "Long before there were people on the earth, crystals were already growing in the earth's crust." \hfill {\scriptsize [Ch.~29]}
\item M.P. Hobson, G.P. Efstathiou, and A.N. Lasenby, \textit{General Relativity: An Introduction for Physicists} (Cambridge: Cambridge University Press, 2006). \hfill {\scriptsize [Ch.~2]}
\item Maimonides (c. 1170-1180). \textit{Mishneh Torah}. \hfill {\scriptsize [Ch.~43]}
\item Maimonides (Moses ben Maimon), \textit{The Guide for the Perplexed} (Moreh Nevukhim), c. 1190. Introduction, section on the definition of faith. Available in S. Pines translation (Chicago: University of Chicago Press, 1963). \hfill {\scriptsize [Ch.~43]}
\item Maki, Z., Nakagawa, M., \& Sakata, S. (1962). Remarks on the unified model of elementary particles. Progress of Theoretical Physics
\item Malachi (~450 BCE). \textit{Book of Malachi}. \hfill {\scriptsize [Ch.~16]}
\item Malachi 3:19 (in Christian numbering 4:1), Hebrew Bible, c. 5th–4th century BCE. \hfill {\scriptsize [Ch.~16]}
\item Malcolm Brown (2005). \textit{The Christmas Truce: The Western Front Comes to a Standstill}. \hfill {\scriptsize [Ch.~38]}
\item Manning, G.S., and Kay, A.R. (2023). The physical basis of osmosis. \textit{J. Gen. Physiol.} 155:e202313332. \hfill {\scriptsize [Ch.~31]}
\item Martin Gardner (1964). \textit{The Ambidextrous Universe: Left, Right, and the Fall of Parity}. Basic Books. \hfill {\scriptsize [Ch.~46]}
\item Martin Gardner (1982). \textit{Aha! Gotcha: Paradoxes to Puzzle and Delight}. W. H. Freeman. \hfill {\scriptsize [Ch.~15]}
\item Masser, D. W. (1985). Lecture on the abc conjecture. Symposium on Diophantine Problems
\item Matiyasevich, Y. (1970). Enumerable sets are Diophantine. Doklady Akademii Nauk SSSR
\item Maurice Kraitchik (1953). \textit{Mathematical Recreations}. Dover Publications. \hfill {\scriptsize [Ch.~15]}
\item Mauss, M. (1925). Essai sur le don. L'Année Sociologique \hfill {\scriptsize [Ch.~38]}
\item Maxwell, J. C. (1860). Illustrations of the dynamical theory of gases. Philosophical Magazine
\item Mayers, D. F. (1957). Relativistic Self-Consistent Field Calculation for Mercury. \textit{Proc. R. Soc. Lond. A}, \textbf{241}, 93. \hfill {\scriptsize [Ch.~1]}
\item Maynard, J. (2015). Small gaps between primes. \textit{Annals of Mathematics}, 181(1), 383–413. \hfill {\scriptsize [Ch.~8]}
\item Meissner, W. \& Ochsenfeld, R. (1933). Ein neuer Effekt bei Eintritt der Supraleitfähigkeit. Naturwissenschaften
\item Meyer, H. (1899). Zur Theorie der Alkoholnarkose. Archiv für experimentelle Pathologie und Pharmakologie
\item Michael Atiyah, "Mathematics in the 20th Century: Geometry versus Algebra," \textit{Mathematics Today} 37, no. 2 (2001), pp. 46–53. \hfill {\scriptsize [Ch.~7]}
\item Milgrom, M. (1983). A modification of the Newtonian dynamics as a possible alternative to the hidden mass hypothesis. The Astrophysical Journal
\item Minkowski, H. (1908). Raum und Zeit. Lecture at the 80th Assembly of German Natural Scientists and Physicians \hfill {\scriptsize [Ch.~6, 47]}
\item Mirollo, R. E. \& Strogatz, S. H. (1990). Synchronization of pulse-coupled biological oscillators. SIAM Journal on Applied Mathematics.
\item Mishnaic sages (c. 200 CE). \textit{Pirkei Avot}. Mishnah. \hfill {\scriptsize [Ch.~26]}
\item Misner, C. W., Thorne, K. S., Wheeler, J. A. (1973). \textit{Gravitation}. W. H. Freeman and Company. \hfill {\scriptsize [Ch.~6, 14]}
\item Mitchell, P. (1961). Coupling of phosphorylation to electron and hydrogen transfer by a chemi-osmotic type of mechanism. Nature.
\item Modris Eksteins (1989). \textit{Rites of Spring: The Great War and the Birth of the Modern Age}. Houghton Mifflin. \hfill {\scriptsize [Ch.~38]}
\item Moore, G. E. (1965). Cramming more components onto integrated circuits. Electronics.
\item Moritz Lipp, Michael Schwarz, Daniel Gruss, Thomas Prescher, Werner Haas, Anders Fogh, Jann Horn, Stefan Mangard, Paul Kocher, Daniel Genkin, Yuval Yarom, Mike Hamburg (2018). \textit{Meltdown: Reading Kernel Memory from User Space}. 27th USENIX Security Symposium. \hfill {\scriptsize [Ch.~18]}
\item Myers, C. S. (1915). A contribution to the study of shell shock. The Lancet.
\item Möbius, A. F. (1865). Über die Bestimmung des Inhaltes eines Polyëders. Berichte über die Verhandlungen der Königlich Sächsischen Gesellschaft der Wissenschaften zu Leipzig.
\item N. D. Birrell and P. C. W. Davies (1982). \textit{Quantum Fields in Curved Space}. Cambridge University Press. \hfill {\scriptsize [Ch.~47]}
\item Nalebuff, B. (1989). The Other Person's Envelope Is Always Greener. \textit{J. Econ. Persp.}, \textbf{3}(1), 171–181. \hfill {\scriptsize [Ch.~15]}
\item Navarro, J. F., Frenk, C. S., and White, S. D. M. (1997). A Universal Density Profile from Hierarchical Clustering. \textit{ApJ}, \textbf{490}, 493–508. \hfill {\scriptsize [Ch.~37]}
\item Newton, I. (1687). Philosophiæ Naturalis Principia Mathematica. Royal Society. \hfill {\scriptsize [Ch.~6, 24]}
\item Nicolas Léonard Sadi Carnot (1824). \textit{Reflections on the Motive Power of Fire}. Bachelier, Paris. \hfill {\scriptsize [Ch.~45]}
\item Niels Bohr, as quoted in Ruth Moore, \textit{Niels Bohr: The Man, His Science, \& the World They Changed} (New York: Knopf, 1966), p. 140. \hfill {\scriptsize [Ch.~15]}
\item Nora, Pierre (1984). \textit{Les Lieux de mémoire}. Gallimard, Paris. \hfill {\scriptsize [Ch.~38]}
\item Norrby, L. J. (1991). Why Is Mercury Liquid? Or, why do relativistic effects not get into chemistry textbooks? \textit{J. Chem. Educ.}, \textbf{68}, 110. \hfill {\scriptsize [Ch.~1]}
\item Oesterlé, J. (1988). Nouvelles approches du "théorème" de Fermat. Séminaire Bourbaki \hfill {\scriptsize [Ch.~33]}
\item Oliver Heaviside (1893). \textit{Electromagnetic Theory}. The Electrician Printing and Publishing Company, London. \hfill {\scriptsize [Ch.~4]}
\item Openshaw, S. (1984). The modifiable areal unit problem. Geo Books.
\item Originated with I.J. Good, address to Institute of Mathematical Statistics, 22 April 1971 (published in \textit{The American Statistician}, June 1972). Ronald H. Coase refined it in a letter dated 17 October 1975: "If you torture the data long enough nature will confess.". \hfill {\scriptsize [Ch.~40]}
\item Osthoff, H. \& Brugmann, K. (1878). Morphologische Untersuchungen auf dem Gebiete der indogermanischen Sprachen. S. Hirzel.
\item Overton, E. (1901). Studien über die Narkose. Gustav Fischer
\item P. W. Kasteleyn (1961). `The statistics of dimers on a lattice: I. The number of dimer arrangements on a quadratic lattice'. \textit{Physica 27, 1209-1225}. \hfill {\scriptsize [Ch.~29]}
\item Padmanabhan, T. (2003). Cosmological constant — the weight of the vacuum. \textit{Phys. Rep.}, 380, 235. \hfill {\scriptsize [Ch.~2]}
\item Palermo, L., von Ranke, P. J., \& da Silva, X. A. (1992). A spin-1/2 Ising model for a semi-infinite lattice in a transverse field. Physical Review B.
\item Pascal, B. (1670). Pensées. Guillaume Desprez.
\item Paul Gordan, as reported in: Max Noether, "Paul Gordan," \textit{Mathematische Annalen} 75 (1914), pp. 1–41. The remark dates to approximately 1888, in response to Hilbert's proof of the finite basis theorem. \hfill {\scriptsize [Ch.~1]}
\item Paul Kocher, Jann Horn, Anders Fogh, Daniel Genkin, Daniel Gruss, Werner Haas, Mike Hamburg, Moritz Lipp, Stefan Mangard, Thomas Prescher, Michael Schwarz, Yuval Yarom (2019). \textit{Spectre Attacks: Exploiting Speculative Execution}. IEEE Symposium on Security and Privacy. \hfill {\scriptsize [Ch.~18]}
\item Pauli, W. (1925). `Über den Zusammenhang des Abschlusses der Elektronengruppen im Atom mit der Komplexstruktur der Spektren'. \textit{Z. Phys.} \textbf{31}, 765–783. \hfill {\scriptsize [Ch.~1]}
\item Pauling, L. (1931). The Nature of the Chemical Bond. Journal of the American Chemical Society.
\item Pavlides, M. G., and Perlman, M. D. (2009). \textit{J. Stat. Plan. Inference}, 139(1), 198–213. \hfill {\scriptsize [Ch.~30]}
\item Penrose, R. (1965). Gravitational Collapse and Space-Time Singularities. Physical Review Letters.
\item Penzias, A. A. \& Wilson, R. W. (1965). A Measurement of Excess Antenna Temperature at 4080 Mc/s. The Astrophysical Journal.
\item Perlin, K. (1985). An Image Synthesizer. ACM SIGGRAPH Computer Graphics.
\item Persson, M. (2009). \textit{Initial Geometry and Entity Source}. Mojang. \hfill {\scriptsize [Ch.~22]}
\item Peter Simkins (1991). \textit{World War I: The Western Front}. \hfill {\scriptsize [Ch.~4]}
\item Pevzner et al. (2001). \textit{PNAS} 98:9748–9753. \hfill {\scriptsize [Ch.~40]}
\item Planck Collaboration (2020). \textit{A\&A}, \textbf{641}, A6. \hfill {\scriptsize [Ch.~43]}
\item Planck, M. (1901). Ueber das Gesetz der Energieverteilung im Normalspectrum. Annalen der Physik.
\item Poincaré, H. (1885). Sur l'équilibre d'une masse fluide animée d'un mouvement de rotation. Acta Mathematica. \hfill {\scriptsize [Ch.~7]}
\item Politzer, H. D. (1973). Reliable Perturbative Results for Strong Interactions?. Physical Review Letters.
\item Pontecorvo, B. (1957). Mesonium and anti-mesonium. Zhurnal Eksperimental'noi i Teoreticheskoi Fiziki.
\item Popper, K. (1934). Logik der Forschung. Julius Springer.
\item Pound, R. V. \& Rebka, G. A. Jr. (1960). Apparent Weight of Photons. Physical Review Letters. \hfill {\scriptsize [Ch.~6, 32]}
\item Poynting, J. H. (1884). On the Transfer of Energy in the Electromagnetic Field. Philosophical Transactions of the Royal Society of London.
\item Prusiner, S. B. (1982). Novel proteinaceous infectious particles cause scrapie. Science.
\item Purcell, E. M., Pound, R. V. (1951). \textit{A Nuclear Spin System at Negative Temperature}. Phys. Rev., 81, 279. \hfill {\scriptsize [Ch.~28]}
\item Putnam, H. (1967). The Nature of Mental States. In Art, Mind, and Religion.
\item Pyykkö, P., Desclaux, J. P. (1979). Relativity and the periodic system of elements. \textit{Acc. Chem. Res.}, \textbf{12} (8), 276–281. \hfill {\scriptsize [Ch.~1]}
\item Rabbi Eliezer ben Hyrcanus (attributed; medieval compilation) (1514). \textit{Pirkei de-Rabbi Eliezer}. Constantinople. \hfill {\scriptsize [Ch.~26]}
\item Rabbi Yisrael Lipschitz, \textit{Tiferet Yisrael} / Boaz on Mishnah Berakhot 1, c. 1830–1850. \hfill {\scriptsize [Ch.~26]}
\item Radon, J. (1917). \"Uber die Bestimmung von Funktionen durch ihre Integralwerte l\"angs gewisser Mannigfaltigkeiten. \textit{Berichte der S\"achsischen Akademie der Wissenschaften zu Leipzig}, 69, 262–277. \hfill {\scriptsize [Ch.~41]}
\item Radó, T. (1962). On non-computable functions. Bell System Technical Journal.
\item Ramsey, N. F. (1956). \textit{Thermodynamics and Statistical Mechanics at Negative Absolute Temperatures}. Phys. Rev., 103, 20. \hfill {\scriptsize [Ch.~28]}
\item Rayleigh, Lord (1871). On the light from the sky, its polarization and color. Philosophical Magazine.
\item René Descartes (1637). \textit{Discours de la méthode}, Part V. Leiden. \hfill {\scriptsize [Ch.~20]}
\item René Descartes (1641). \textit{Meditations on First Philosophy}. Michel Soly. \hfill {\scriptsize [Ch.~50]}
\item René Descartes, \textit{Principia Philosophiae} (Amsterdam: Elsevier, 1644), Part II. \hfill {\scriptsize [Ch.~47]}
\item Richard Nixon — Statement made in 1972 regarding inflation rates. Widely reported. \hfill {\scriptsize [Ch.~13, 21]}
\item Richard O. Duda, Peter E. Hart (1973). \textit{Pattern Classification and Scene Analysis}. Wiley-Interscience. \hfill {\scriptsize [Ch.~13]}
\item Richard P. Feynman (1985). \textit{QED: The Strange Theory of Light and Matter}. Princeton University Press. \hfill {\scriptsize [Ch.~13]}
\item Riemann, B. (1859). Ueber die Anzahl der Primzahlen unter einer gegebenen Grösse. Monatsberichte der Berliner Akademie. \hfill {\scriptsize [Ch.~8]}
\item Riess, A. G. et al. (1998). Observational Evidence from Supernovae for an Accelerating Universe and a Cosmological Constant. \textit{Astron. J.} \textbf{116}, 1009–1038. \hfill {\scriptsize [Ch.~2, 34, 43]}
\item Ringe, D. (2006). \textit{From Proto-Indo-European to Proto-Germanic}. Oxford University Press. \hfill {\scriptsize [Ch.~5]}
\item Robert L. Berger (1966). \textit{The Undecidability of the Domino Problem}. Memoirs of the American Mathematical Society. \hfill {\scriptsize [Ch.~29]}
\item Robert M. Wald (1984). \textit{General Relativity}. University of Chicago Press. \hfill {\scriptsize [Ch.~23]}
\item Robinson, J. (1950). Existential definability in arithmetic. Transactions of the American Mathematical Society. \hfill {\scriptsize [Ch.~30]}
\item Robinson, R. M. (1971). \textit{Undecidability and Nonperiodicity for Tilings of the Plane}. Invent. Math. \hfill {\scriptsize [Ch.~29]}
\item Roger Penrose (2004). \textit{The Road to Reality: A Complete Guide to the Laws of the Universe}. Jonathan Cape. \hfill {\scriptsize [Ch.~29, 34]}
\item Rossi, B., and Hall, D. B., \textit{Physical Review}, 59(3), 223–228 (1941). \hfill {\scriptsize [Ch.~19]}
\item Royen, T. (2014). \textit{A simple proof of the Gaussian correlation conjecture extended to multivariate gamma distributions}. Far East J. Theor. Stat. \hfill {\scriptsize [Ch.~33]}
\item Ruben, S., Randall, M., Kamen, M., and Hyde, J. L. (1941). Heavy Oxygen ($^{18}$O) as a Tracer in the Study of Photosynthesis. \textit{J. Am. Chem. Soc.} \textbf{63}, 877–879. \hfill {\scriptsize [Ch.~35]}
\item Rybicki, G. B., and Lightman, A. P. (1979). \textit{Radiative Processes in Astrophysics}. Wiley. \hfill {\scriptsize [Ch.~27]}
\item S Coleman and F De Luccia (1980). `Gravitational effects on and of vacuum decay'. \textit{Physical Review D}. \hfill {\scriptsize [Ch.~16]}
\item S.W. Hawking (1974). `Black hole explosions?'. \textit{Nature 248, 30-31}. \hfill {\scriptsize [Ch.~23]}
\item Safronova, M. S. et al. (2018). \textit{Search for new physics with atomic clocks}. Rev. Mod. Phys. 90(2), 025008. \hfill {\scriptsize [Ch.~32]}
\item Sagnac, G. (1913). L'éther lumineux démontré par l'effet du vent relatif d'éther dans un interféromètre en rotation uniforme. Comptes Rendus de l'Académie des Sciences.
\item Sanger, F., Nicklen, S., \& Coulson, A. R. (1977). DNA sequencing with chain-terminating inhibitors. Proceedings of the National Academy of Sciences.
\item Sara Lewis (2016). \textit{Silent Sparks: The Wondrous World of Fireflies}. Princeton University Press. \hfill {\scriptsize [Ch.~25]}
\item Satterthwaite, M. A. (1975). Strategy-proofness and Arrow's conditions. \textit{J. Econ. Theory} \textbf{10}(2), 187–217. \hfill {\scriptsize [Ch.~9]}
\item Schiffer, P. H. et al. (2018). The nematode \textit{Acrobeloides nanus} reveals conserved phylum-specific expression patterns. \textit{PNAS} \textbf{115}(6), 1269–1274. \hfill {\scriptsize [Ch.~40]}
\item Schleicher, A. (1861). Compendium der vergleichenden Grammatik der indogermanischen Sprachen. H. Böhlau.
\item Schrödinger, E. (1926). `Quantisierung als Eigenwertproblem'. \textit{Annalen der Physik} \textbf{384}(4), 361–376. \hfill {\scriptsize [Ch.~1]}
\item Schwarzschild, K. (1916). Über das Gravitationsfeld eines Massenpunktes nach der Einsteinschen Theorie. Sitzungsberichte der Königlich Preussischen Akademie der Wissenschaften.
\item Sean M. Carroll (2004). \textit{Spacetime and Geometry: An Introduction to General Relativity}. Addison Wesley. \hfill {\scriptsize [Ch.~2]}
\item Searle, J. R. (1980). \textit{Minds, Brains, and Programs}. Behavioral and Brain Sciences, 3(3), 417–457. \hfill {\scriptsize [Ch.~1, 20]}
\item Shannon, C. E. (1938). A Symbolic Analysis of Relay and Switching Circuits. Transactions of the American Institute of Electrical Engineers. \hfill {\scriptsize [Ch.~22]}
\item Shannon, C. E. \& Weaver, W. (1949). The Mathematical Theory of Communication. University of Illinois Press.
\item Shechtman, D., Blech, I., Gratias, D., and Cahn, J. W. (1984). Metallic Phase with Long-Range Orientational Order and No Translational Symmetry. \textit{Phys. Rev. Lett.} \textbf{53}, 1951–1953. \hfill {\scriptsize [Ch.~29]}
\item Silver, D. H. et al. (2013). ELOPER: elongation of paired-end reads. \textit{Bioinformatics} \textbf{29}(11), 1465–1467. \hfill {\scriptsize [Ch.~40]}
\item Simpson, E. H. (1951). \textit{J. R. Stat. Soc. B}, 13(2), 238–241. \hfill {\scriptsize [Ch.~30]}
\item Slater, B., and Michaelides, A. (2019). \textit{Nat. Rev. Chem.}, \textbf{3}, 172. \hfill {\scriptsize [Ch.~42]}
\item Sobel, I. (1970). Camera Models and Machine Perception. Stanford University.
\item Soddy, F. (1913). The Radio-Elements and the Periodic Law. Nature.
\item Sonnino, Paul (2016). \textit{The Search for the Man in the Iron Mask: A Historical Detective Story}. Rowman \& Littlefield. \hfill {\scriptsize [Ch.~39, 44]}
\item Sophie Germain (1821). \textit{Recherches sur la théorie des surfaces élastiques}. Académie des sciences. \hfill {\scriptsize [Ch.~13]}
\item Spergel, D. N. et al. (2003). First-Year WMAP Observations. \textit{ApJS} \textbf{148}, 175–194. \hfill {\scriptsize [Ch.~37]}
\item Steinhardt, P. J. \& Turok, N. (2002). A Cyclic Model of the Universe. Science.
\item Stephanopoulos, N. O. \& McGhee, E. M. (2015). Partisan Gerrymandering and the Efficiency Gap. University of Chicago Law Review.
\item Stern, Sacha (2019). \textit{The Jewish Calendar Controversy of 921/2 CE}. Brill. \hfill {\scriptsize [Ch.~26]}
\item Swadesh, M. (1955). Towards greater accuracy in lexicostatistic dating. International Journal of American Linguistics.
\item Szilard, L. (1929). \textit{On the Decrease of Entropy in a Thermodynamic System by the Intervention of Intelligent Beings}. Z. Phys., 53, 840–856. \hfill {\scriptsize [Ch.~45]}
\item Tachibana, M. \textit{et al.} (2009). Mitochondrial gene replacement in primate offspring and embryonic stem cells. \textit{Nature} \textbf{461}, 367–372. \hfill {\scriptsize [Ch.~49]}
\item Tachibana, M. \textit{et al.} (2013). Towards germline gene therapy of inherited mitochondrial diseases. \textit{Nature} \textbf{493}, 627–631. \hfill {\scriptsize [Ch.~49]}
\item Tanya Lee Stone (2003). \textit{Ilan Ramon: Israel's First Astronaut}. Millbrook Press. \hfill {\scriptsize [Ch.~44]}
\item Terraine, J. (1978). \textit{To Win a War: 1918, the Year of Victory}. Sidgwick \& Jackson. \hfill {\scriptsize [Ch.~38]}
\item The National Archives (UK) (1914). `BEF Unit War Diaries, Dec 1914'. \textit{The National Archives (UK)}. \hfill {\scriptsize [Ch.~38]}
\item Thomas Reid (1785). \textit{Essays on the Intellectual Powers of Man}. J. Bell, Edinburgh. \hfill {\scriptsize [Ch.~50]}
\item Thouless, D. J., Kohmoto, M., Nightingale, M. P., and den Nijs, M. (1982). Quantized Hall Conductance in a Two-Dimensional Periodic Potential. \textit{Phys. Rev. Lett.} \textbf{49}, 405–408. \hfill {\scriptsize [Ch.~11]}
\item Timmins, G.S., et al. (2001). Firefly flashing is controlled by gating oxygen to light-emitting cells. \textit{J. Exp. Biol.}, 204, 2795–2801. \hfill {\scriptsize [Ch.~25]}
\item Tomasulo, R. M. (1967). An Efficient Algorithm for Exploiting Multiple Arithmetic Units. IBM Journal of Research and Development.
\item Tononi, G. (2004). An information integration theory of consciousness. BMC Neuroscience.
\item Turing, A. M. (1936). On Computable Numbers, with an Application to the Entscheidungsproblem. Proceedings of the London Mathematical Society. \hfill {\scriptsize [Ch.~17]}
\item Turing, A. M. (1950). Computing Machinery and Intelligence. Mind. \hfill {\scriptsize [Ch.~20]}
\item Unruh, W. G. (1976). Notes on black-hole evaporation. Physical Review D.
\item van Ginkel, M., Luengo Hendriks, C. L., and van Vliet, L. J. (2004). \textit{A short introduction to the Radon and Hough transforms and how they relate to each other}. Imaging Science and Technology Department, Delft University of Technology, The Netherlands. \hfill {\scriptsize [Ch.~41]}
\item Various (Tannaitic sages, school of Rabbi Akiva) (c. 200 CE). \textit{Midrash Halakha}. \hfill {\scriptsize [Ch.~5]}
\item Vegard, L. (1908). Zur Theorie der osmotischen Erscheinungen. \textit{Proc. Camb. Phil. Soc.} 15:13–23. \hfill {\scriptsize [Ch.~31]}
\item Vernam, G. S. (1926). Cipher Printing Telegraph Systems For Secret Wire and Radio Telegraphic Communications. Journal of the American Institute of Electrical Engineers.
\item Voltaire (1751). \textit{Le Siècle de Louis XIV}. C. F. Henning. \hfill {\scriptsize [Ch.~39, 44]}
\item von Neumann, John, and Morgenstern, Oskar (1944). \textit{Theory of Games and Economic Behavior}. Princeton University Press. \hfill {\scriptsize [Ch.~3]}
\item W. Deane Butcher, Translator's Preface to Stéphane Leduc, \textit{The Mechanism of Life} (London: William Heinemann, 1911), p. vii. \hfill {\scriptsize [Ch.~31]}
\item Wallace, D. C., et al. (1988). Mitochondrial DNA mutation associated with Leber's hereditary optic neuropathy. Science.
\item Wang, H. (1961). Proving theorems by pattern recognition — II. Bell System Technical Journal.
\item Weber, B. et al. (2018). \textit{J. Phys. Chem. Lett.}, \textbf{9}, 2838. \hfill {\scriptsize [Ch.~42]}
\item Weinberg, S. (1979). Ultraviolet divergences in quantum theories of gravitation. In S. W. Hawking and W. Israel (Eds.), \textit{General Relativity: An Einstein Centenary Survey} (Cambridge University Press). \hfill {\scriptsize [Ch.~36]}
\item Weinberg, S. (1989). The cosmological constant problem. \textit{Rev. Mod. Phys.}, 61, 1. \hfill {\scriptsize [Ch.~2]}
\item Weinberg, Steven (1995). \textit{The Quantum Theory of Fields, Vol. I: Foundations}. Cambridge University Press. \hfill {\scriptsize [Ch.~24]}
\item Weintraub, Stanley (2001). \textit{Silent Night: The Story of the World War I Christmas Truce}. Free Press. \hfill {\scriptsize [Ch.~38]}
\item Weyl, H. (1916). Über die Gleichverteilung von Zahlen mod. Eins. Mathematische Annalen. \hfill {\scriptsize [Ch.~7]}
\item Wiedemann, G., \& Franz, R. (1853). Ueber die Wärme-Leitungsfähigkeit der Metalle. Annalen der Physik.
\item Wien, W. (1896). Ueber die Energieverteilung im Emissionsspectrum eines schwarzen Körpers. Annalen der Physik.
\item Wiles, A. (1995). Modular elliptic curves and Fermat's Last Theorem. Annals of Mathematics.
\item Williams, A. O. (1940). A Relativistic Self-Consistent Field for Cu\(^+\). \textit{Phys. Rev.}, \textbf{58}, 723. \hfill {\scriptsize [Ch.~1]}
\item Wills, C. (1967). The ecological relations of California voles. Ecology.
\item Wilson, K. G. (1971). Renormalization Group and Critical Phenomena. I. Renormalization Group and the Kadanoff Scaling Picture. Physical Review B.
\item Winston Churchill, speech to the House of Commons, 11 November 1947. Published in \textit{Hansard}, vol. 444, col. 206–7. \hfill {\scriptsize [Ch.~9]}
\item Woodward, R. B., and Hoffmann, R. (1970). \textit{The Conservation of Orbital Symmetry}. Verlag Chemie GmbH / Academic Press. \hfill {\scriptsize [Ch.~46]}
\item Wu, A. (2008). `The Banach-Tarski Paradox'. \textit{University of Chicago VIGRE REU}. \hfill {\scriptsize [Ch.~3]}
\item Yang, C. N., \& Mills, R. L. (1954). Conservation of Isotopic Spin and Isotopic Gauge Invariance. Physical Review.
\item Yehudai Gaon (8th century). \textit{Halakhot Pesukot}. \hfill {\scriptsize [Ch.~26]}
\item Zermelo, E. (1904). Beweis, daß jede Menge wohlgeordnet werden kann. Mathematische Annalen. \hfill {\scriptsize [Ch.~3, 7]}
\item Zhang, Y. (2014). Bounded gaps between primes. \textit{Annals of Mathematics}, 179(3), 1121–1174. \hfill {\scriptsize [Ch.~8]}
\item Zimbardo, P. G. (1971). The Power and Pathology of Imprisonment. Congressional Record.
\item Zwicky, F. (1933). Die Rotverschiebung von extragalaktischen Nebeln. \textit{Helv. Phys. Acta}, \textbf{6}, 110–127. \hfill {\scriptsize [Ch.~37]}
\end{list}

\subsection*{Literary, Film \& Media Sources}
\vspace{0.3em}
\begin{list}{}{%
  \setlength{\leftmargin}{2em}%
  \setlength{\itemindent}{-2em}%
  \setlength{\itemsep}{0.35\baselineskip}%
  \setlength{\parsep}{0pt}%
  \setlength{\topsep}{0.5em}}
\small
\item "Let the Lord of Chaos rule." — Dark One's command, \textit{Lord of Chaos} (1994) Prologue. Semirhage says it in \textit{Knife of Dreams} (2005), Ch. 34. Robert Jordan, \textit{The Wheel of Time}. \hfill {\scriptsize [Ch.~48]}
\item "The Gang" — \textit{It's Always Sunny in Philadelphia}, Season 7, Episode 4, "Sweet Dee Gets Audited" (FX, 2011). \hfill {\scriptsize [Ch.~50]}
\item Albert Einstein — "Zeit ist das, was man an der Uhr abliest." From a 1926 discussion with Rudolf Carnap, as reported in Carnap's autobiography in \textit{The Philosophy of Rudolf Carnap} (La Salle: Open Court, 1963), p. 37. \hfill {\scriptsize [Ch.~6]}
\item Albert Einstein, letter to Cornel Lanczos, 21 March 1942. Published in \textit{Albert Einstein: The Human Side} (Princeton: Princeton University Press, 1979), eds. Helen Dukas and Banesh Hoffmann. \hfill {\scriptsize [Ch.~6]}
\item Andy Dwyer — \textit{Parks and Recreation}, Season 6, Episode 1, "London, Part 1" (NBC, 2013). Fictional character played by Chris Pratt. \hfill {\scriptsize [Ch.~42]}
\item Brandon Sanderson (2010). \textit{The Way of Kings (The Stormlight Archive, Book 1)}. Tor Books. \hfill {\scriptsize [Ch.~18]}
\item Charlie Kelly — \textit{It's Always Sunny in Philadelphia}, Season 5, Episode 3, "The Great Recession" (FX, 2009). \hfill {\scriptsize [Epigraph, Ch.~27]}
\item Cohen the Barbarian — Terry Pratchett, \textit{The Last Hero} (London: Gollancz, 2001). Fictional character. Quote also used by @terryandrob to announce Pratchett's death, 12 March 2015. \hfill {\scriptsize [Ch.~3]}
\item Comte de Lautréamont (1869). \textit{Les Chants de Maldoror}. Lacroix, Verboeckhoven et Cie. \hfill {\scriptsize [Ch.~25]}
\item Dan Simmons (1989). \textit{Hyperion}. Doubleday. \hfill {\scriptsize [Ch.~20]}
\item Dave Barry, \textit{Dave Barry Turns 50} (New York: Crown, 1998). \hfill {\scriptsize [Ch.~4]}
\item Douglas Adams, \textit{The Hitchhiker's Guide to the Galaxy} (London: Pan Books, 1979), Chapter 1. \hfill {\scriptsize [Ch.~14]}
\item Dr. Seuss (Theodor Seuss Geisel), \textit{The Lorax} (New York: Random House, 1971). \hfill {\scriptsize [Ch.~35]}
\item Eddard Stark — George R.R. Martin, \textit{A Game of Thrones} (New York: Bantam, 1996), Chapter 1 ("Bran I"). \hfill {\scriptsize [Ch.~28]}
\item George R. R. Martin (1996). \textit{A Game of Thrones}. Bantam Books. \hfill {\scriptsize [Ch.~28]}
\item Greg Egan (1994). \textit{Permutation City}. Millennium Orion Publishing Group. \hfill {\scriptsize [Ch.~39]}
\item I.I. Rabi — "Who ordered that?" upon learning of the muon's discovery, 1936. Widely reported; see e.g., Robert P. Crease and Charles C. Mann, \textit{The Second Creation} (New York: Macmillan, 1986). \hfill {\scriptsize [Ch.~19]}
\item Jack Donaghy — \textit{30 Rock}, Season 1, Episode 19, "Corporate Crush" (NBC, 2007). Fictional character played by Alec Baldwin. \hfill {\scriptsize [Ch.~14]}
\item J.R.R. Tolkien, \textit{The Fellowship of the Ring} (London: George Allen \& Unwin, 1954), poem in Chapter 10 ("Strider"). \hfill {\scriptsize [Ch.~1]}
\item Jacob Grimm, Wilhelm Grimm (1812). \textit{Kinder- und Hausmärchen}. Realschulbuchhandlung, Berlin. \hfill {\scriptsize [Ch.~7]}
\item Joanna Newsom, "Cosmia," \textit{Ys} (Drag City, 2006). Song lyrics. \hfill {\scriptsize [Ch.~16]}
\item Kenneth Parcell — \textit{30 Rock}, Season 5, Episode 19, "I Heart Connecticut" (NBC, 2011). \hfill {\scriptsize [Ch.~6]}
\item Lewis Carroll, \textit{Alice's Adventures in Wonderland} (London: Macmillan, 1865), Chapter 2. \hfill {\scriptsize [Ch.~25]}
\item Logen Ninefingers — Joe Abercrombie, \textit{The Blade Itself} (London: Gollancz, 2006). \hfill {\scriptsize [Ch.~22]}
\item Mac — \textit{It's Always Sunny in Philadelphia}, Season 5, Episode 3, "The Great Recession" (FX, 2009). "That doesn't sound right, but I don't know enough about stars to dispute it." \hfill {\scriptsize [Ch.~15]}
\item Mark Twain, \textit{Following the Equator} (Hartford: American Publishing Company, 1897), Chapter 15 epigraph. \hfill {\scriptsize [Ch.~3]}
\item Martin Silenus — Dan Simmons, \textit{Hyperion} (New York: Doubleday, 1989), "The Poet's Tale.". \hfill {\scriptsize [Ch.~20]}
\item Nicomo Cosca — Joe Abercrombie, \textit{Best Served Cold} (London: Gollancz, 2009). \hfill {\scriptsize [Ch.~38]}
\item Solid Jackson — Terry Pratchett, \textit{Jingo} (London: Gollancz, 1997). \hfill {\scriptsize [Ch.~28]}
\item Super Hans — \textit{Peep Show}, Series 3, Episode 2, "Sectioning" (Channel 4, 2005). Fictional character played by Matt King. \hfill {\scriptsize [Ch.~9]}
\item Terry Pratchett (1997). \textit{Jingo}. Victor Gollancz. \hfill {\scriptsize [Ch.~7]}
\item The Dogman — Joe Abercrombie, \textit{Before They Are Hanged} (London: Gollancz, 2007). Fictional character. \hfill {\scriptsize [Ch.~43]}
\item The Weakerthans, "Left and Leaving," \textit{Left and Leaving} (G7 Welcoming Committee Records, 2000). Song lyrics by John K. Samson. \hfill {\scriptsize [Ch.~44]}
\item Tiffany Aching — Terry Pratchett, \textit{A Hat Full of Sky} (London: Doubleday, 2004). \hfill {\scriptsize [Ch.~46]}
\item Tobias Funke — \textit{Arrested Development}, Season 2, Episode 5, "Sad Sack" (Fox, 2004). \hfill {\scriptsize [Ch.~12]}
\item Tom Cargill — "Ninety-ninety rule." Reported in Jon Bentley, "Programming Pearls," \textit{Communications of the ACM} 28, no. 9 (September 1985). \hfill {\scriptsize [Ch.~22]}
\item Wit (Hoid) to Shallan — Brandon Sanderson, \textit{Oathbringer} (New York: Tor Books, 2017), Chapter 68, "Aim for the Sun." Fictional character. \hfill {\scriptsize [Ch.~18]}
\end{list}
\vspace{1em}

\endgroup