\begin{technical}
  \textbf{\Large Poncelet for Two Ellipses}\vspace{0.2em}
  
  Let \( C \) and \( D \) be two smooth, strictly convex, nested ellipses in the plane, with \( D \subset \operatorname{int}(C) \). The \emph{Poncelet map} \( T: C \to C \) is defined as follows: for a point \( p \in C \), draw the line through \( p \) tangent to the inner ellipse \( D \) (choosing one of the two tangent directions consistently and smoothly along \( C \)); let \( T(p) \) be the second point of intersection of this line with \( C \). This consistent choice defines an orientation-preserving homeomorphism of \( C \).
  
  We show that \( T \) preserves a natural measure and is topologically conjugate to a circle rotation. This yields a complete classification of the dynamics of \( T \), and with it, a proof of Poncelet's closure theorem.
  
  \textbf{Affine Reduction and the Invariant Measure}
  To analyze \( T \), we apply two affine reductions. First, by dilating along one axis of \( D \), we may assume the inner ellipse \( D \) is a circle; this preserves incidence and tangency, so the Poncelet map is unchanged. Now suppose \( C \) is given by
  \[
  {x^2}/{a^2} + {y^2}/{b^2} = 1, \quad \text{with } a > b > 0.
  \]
  Let \( U(x, y) = (x/a, y/b) \), sending \( C \) to the unit circle \( \widetilde{C} \) and \( D \) to an ellipse \( \widetilde{D} \). Let \( \widetilde{s} \) denote arc-length on \( \widetilde{C} \). For \( p \in C \), let \( m \) be the tangency point of the chord \( p\,T(p) \) with \( D \), and let \( m_1 \) be the tangency point of the next chord \( T(p)\,T^2(p) \). Set \( \rho(p) := |p - m| \). The invariant measure is \( d\mu := d\widetilde{s}\,/\,\rho \), mixing arc-length on \( \widetilde{C} \) with the tangent-segment distance in the original (circular-\(D\)) coordinates.
  
  \textbf{Invariance of the Measure.} For nearby \( \widetilde{p}, \widetilde{p}' \in \widetilde{C} \), the chords \( \widetilde{p}\widetilde{p}_1 \) and \( \widetilde{p}'\widetilde{p}_1' \) meet at a point \( \widetilde{n} \), forming similar triangles. As \( \widetilde{p}' \to \widetilde{p} \), \( \widetilde{n} \to \widetilde{m} \) and chord ratios on \( \widetilde{C} \) converge to arc-length ratios (\( \sin\theta/\theta \to 1 \)), so
  \[
  \frac{d\widetilde{s}_1}{d\widetilde{s}} = \frac{|\widetilde{p}_1\widetilde{m}|}{|\widetilde{p}\widetilde{m}|} = \frac{|p_1 m|}{|pm|},
  \]
  where the second equality holds because \( U \) is linear and preserves distance ratios. Since \( D \) is a circle, the two tangent segments from \( p_1 \) to \( D \) are equal: \( |p_1 m| = |p_1 m_1| = \rho(T(p)) \). Hence
  \[
  \frac{d\widetilde{s}_1}{d\widetilde{s}} = \frac{\rho(T(p))}{\rho(p)}, \quad \text{i.e.,} \quad \frac{d\widetilde{s}}{\rho(p)} = \frac{d\widetilde{s}_1}{\rho(T(p))},
  \]
  and \( d\mu = d\widetilde{s}\,/\,\rho \) is \( T \)-invariant.
  
  \textbf{Topological Conjugacy to a Circle Rotation}
  An orientation-preserving circle homeomorphism with a finite non-atomic invariant measure positive on every non-empty open arc has no wandering intervals (a result due to Poincaré); therefore its Poincaré semiconjugacy to a rotation is a conjugacy.
  
  Our measure \( \mu \) satisfies: \textbf{Finite:} Since \( C \) and \( D \) are smooth and strictly convex with \( D \subset \operatorname{int}(C) \), \( \tilde{\rho} \) is continuous on the compact set \( C' \) and strictly positive, so it has a positive infimum. Thus \( \int_{C'} d\tilde{s}/\tilde{\rho} < \infty \). \textbf{Non-atomic:} points have zero measure. \textbf{Positive on arcs:} \( \tilde{\rho} \) is bounded above on any arc, so \( 1/\tilde{\rho} \) is bounded below by a positive constant, ensuring open arcs have positive measure.
  
  Thus, there exists a homeomorphism \( \varphi: C \to S^1 \) and \( \alpha \in [0,1) \) such that: $ \varphi \circ T \circ \varphi^{-1} = R_\alpha, \quad \text{where } R_\alpha(\theta) = \theta + \alpha \mod 1$
  
  The number \( \alpha \), called the \emph{rotation number} of \( T \), quantifies the average angular displacement per iteration. Lift the circle map to \( F: \mathbb{R} \to \mathbb{R} \) in an angular coordinate; then $\alpha := \lim_{n \to \infty} {(F^n(\theta) - \theta)}/{n}$ exists and is independent of \( \theta \). The rotation number of \( T \) is \( \alpha \bmod 1 \).
  
  The rotation number classifies the dynamics:
  \begin{itemize}
    \item If \( \alpha \in \mathbb{Q} \), write \( \alpha = {m}/{n} \) in lowest terms. Then all orbits are periodic with period \( n \). Every point on \( C \) traces a closed \( n \)-gon tangent to \( D \).
    \item If \( \alpha \notin \mathbb{Q} \), then no orbit is periodic. The sequence \( p, T(p), T^2(p), \ldots \) becomes dense in \( C \), and no Poncelet polygon closes.
  \end{itemize}
  
  \techref
  {\footnotesize
  Leopold Flatto, \textit{Poncelet's Theorem}. Mathematical Surveys and Monographs, Vol. 56. American Mathematical Society, 2009 (beautiful book!)\\
  For visualization, see \href{https://bit.ly/poporism}{\textcolor{black}{bit.ly/poporism}}.\\
  }
  \end{technical}
  