\begin{SideNotePage}{
  \textbf{Hilbert's Hotel with Infinite Buses:} An infinite number of buses, each carrying infinitely many passengers, arrives at a hotel with countably infinite rooms. To accommodate everyone, each passenger \(n\) on bus \(m\) is assigned to room \((p_n)^m\), where \(p_n\) is the \(n\)th prime number. No two passengers ever collide, since distinct prime powers are never equal: for \(p \ne q\), there are no integers \(k, \ell\) such that \(p^k = q^\ell\). \par

  \textbf{Banach-Tarski as a Card Game:} Imagine a deck with cards \(A, B, A^{-1}, B^{-1}\), where you can make any pile of cards, under one rule: $A$ and $A^{-1}$ cancel, $B$ and $B^{-1}$ cancel (the cards disappear if they are next to one another). Each card represents a rigid rotation: \(A, A^{-1}\) rotate around one axis; \(B, B^{-1}\) around another. In the diagram, blue and red are inverses, as are green and yellow.

  Four players each take all reduced sequences (no cancelling cards neighbour each other) beginning with a specific card, forming four disjoint point sets: \(S_A, S_{A^{-1}}, S_B, S_{B^{-1}}\).

  Now prepend an \(A\) to every sequence in \(S_{A^{-1}}\), practically rotating the whole set, forming \(A \cdot S_{A^{-1}}\). This transforms sequences that began with \(A^{-1}\) into sequences that no longer do. For instance, \(A^{-1}BA \mapsto BA\), and \(A^{-1} \mapsto\) identity. The union \(S_A \cup A \cdot S_{A^{-1}}\) now covers all sequences—forming one full copy of the sphere. Similarly, \(S_B \cup B \cdot S_{B^{-1}}\) forms a second. \par
}{03_BanachTarskiParadox/01_ An Axiom of Your Choice.pdf}
\end{SideNotePage}
